---
title: 'Bios 6301: Assignment 3'
author: "Lan Shi"
output:
  pdf_document: default
  html_document: default
---

*Due Tuesday, 28 September, 1:00 PM*

50 points total.

Add your name as `author` to the file's metadata section.

Submit a single knitr file (named `homework3.rmd`) by email to michael.l.williams@vanderbilt.edu.
Place your R code in between the appropriate chunks for each question.
Check your output by using the `Knit HTML` button in RStudio.

$5^{n=day}$ points taken off for each day late.

### Question 1 ###

**15 points**

Write a simulation to calculate the power for the following study
design.  The study has two variables, treatment group and outcome.
There are two treatment groups (0, 1) and they should be assigned
randomly with equal probability.  The outcome should be a random normal
variable with a mean of 60 and standard deviation of 20.  If a patient
is in the treatment group, add 5 to the outcome.  5 is the true
treatment effect.  Create a linear model for the outcome by the
treatment group, and extract the p-value (hint: see assigment1).
Test if the p-value is less than or equal to the alpha level, which
should be set to 0.05.

Repeat this procedure 1000 times. The power is calculated by finding
the percentage of times the p-value is less than or equal to the alpha
level.  Use the `set.seed` command so that the professor can reproduce
your results.

1. Find the power when the sample size is 100 patients. (10 points)

```{r}
set.seed(925)
n = 100
mean(replicate(1e3, {
    treat_grp = rbinom(n, 1, 0.5)  # treat_grp = 1, if in treatment group
    outcome = rnorm(n, mean=60, sd=20)
    outcome[treat_grp==1] = outcome[treat_grp==1] + 5
    dt = data.frame(treat_grp,outcome)
    fit = lm(outcome~treat_grp,dt)
    #pvalue 
    coef(summary(fit))[2,4]}) < 0.05)
```

1. Find the power when the sample size is 1000 patients. (5 points)

```{r}
set.seed(925)
n = 1000
mean(replicate(1e3, {
    treat_grp = rbinom(n, 1, 0.5)  # treat_grp = 1, if in treatment group
    outcome = rnorm(n, mean=60, sd=20)
    outcome[treat_grp==1] = outcome[treat_grp==1] + 5
    dt = data.frame(treat_grp,outcome)
    fit = lm(outcome~treat_grp,dt)
    #pvalue 
    coef(summary(fit))[2,4]}) < 0.05)
```

### Question 2 ###

**14 points**

Obtain a copy of the [football-values lecture](https://github.com/couthcommander/football-values).
Save the `2021/proj_wr21.csv` file in your working directory.  Read
in the data set and remove the first two columns.

```{r}
dt = read.csv('~/Desktop/21 FA/6301_Stats_Computing/Bios6301-main/football-values-main/2021/proj_wr21.csv')[,-(1:2)]
```

1. Show the correlation matrix of this data set. (4 points)

```{r}
(rho.dt=cor(dt)) # correlation
```

1. Generate a data set with 30 rows that has a similar correlation
structure.  Repeat the procedure 1,000 times and return the mean
correlation matrix. (10 points)

```{r}
# codes are cited from lecture 9
library(MASS)
set.seed(925)

# Assume the joint distribution is normal 
means.dt = colMeans(dt)
vcov.dt = var(dt)
(rho.dt = cor(dt))

# Generate a data set with similar correlation structure
dt.sim = mvrnorm(30, mu = means.dt, Sigma = vcov.dt)
cor(dt.sim)

# repeat 1000 times.
rho.sim = 0
loops=1e3

for (i in 1:loops) {
      dt.sim = mvrnorm(30, mu = means.dt, Sigma = vcov.dt)
      rho.sim = rho.sim+cor(dt.sim)/loops
}
# mean correlation matrix
rho.sim
```

### Question 3 ###

**21 points**

Here's some code:

```{r}
nDist <- function(n = 100) {
    df <- 10
    prob <- 1/3
    shape <- 1
    size <- 16
    list(
        beta = rbeta(n, shape1 = 5, shape2 = 45),
        binomial = rbinom(n, size, prob),
        chisquared = rchisq(n, df),
        exponential = rexp(n),
        f = rf(n, df1 = 11, df2 = 17),
        gamma = rgamma(n, shape),
        geometric = rgeom(n, prob),
        hypergeometric = rhyper(n, m = 50, n = 100, k = 8),
        lognormal = rlnorm(n),
        negbinomial = rnbinom(n, size, prob),
        normal = rnorm(n),
        poisson = rpois(n, lambda = 25),
        t = rt(n, df),
        uniform = runif(n),
        weibull = rweibull(n, shape)
    )
}
```

1. What does this do? (3 points)

    ```{r}
    round(sapply(nDist(500), mean), 2)
    ```
    
    ```
    It gives the mean of the 500 random generated samples from each distrition, and rounds off these values to two decimal places.
    ```

1. What about this? (3 points)

    ```{r}
    sort(apply(replicate(20, round(sapply(nDist(10000), mean), 2)), 1, sd))
    ```
    
    ```
    First, for each distribution, it randomly sampled 10000 samples and calculated their means (w/ two decimal places), and repeated this process for 20 times; Then it calculated the standard deviation of the 20 means of each distribution, and arrange these standard deviations in an numerically increasing order.
    ```

    In the output above, a small value would indicate that `N=10,000` would provide a sufficent sample size as to estimate the mean of the distribution. Let's say that a value *less than 0.02* is "close enough".

1. For each distribution, estimate the sample size required to simulate the distribution's mean. (15 points)

```{r, eval = FALSE}
# suff_size: used to store sufficient sample size for each distribution.
suff_size = numeric(15)
suff_sd = numeric(15) 
names(suff_size) = names(sapply(nDist(1000), mean)) 
names(suff_sd) = names(sapply(nDist(1000), mean)) 
#dist: indicator of distribution
# 1 for beta, 2 for binomial, ..., 15 for weibull
set.seed(925)
for (dist in 1:15){  
    n = 0
    std_dev = Inf
    if (dist %in% c(1,14)){ # use different increments for different distribution.
        gap = 5
    }else if(dist %in% c(5,15,11,4,8,6)){
        gap = 200
    }else if(dist %in% c(13,2,7)){
        gap = 1000
    }else{
        gap = 3000
    }
    while (std_dev>=.02){
        n = n+ gap
        sample_sd = apply(replicate(20, round(sapply(nDist(n), mean), 2)), 1, sd) 
        std_dev = sample_sd[dist] 
    }
    suff_size[dist] = n
    suff_sd[dist] = std_dev
}
suff_size
suff_sd
```

Don't worry about being exact. It should already be clear that N < 10,000 for many of the distributions. You don't have to show your work. Put your answer to the right of the vertical bars (`|`) below.

distribution|N
---|---
beta|5
binomial|7000
chisquared|33000
exponential|3400
f|1000
gamma|1400
geometric|13000
hypergeometric|3800
lognormal|9000
negbinomial|165000 
normal|1800  
poisson|57000
t|3000
uniform|160
weibull|2400
