---
title: 'Bios 6301: Assignment 7'
output: pdf_document
author: 'Lan Shi'
editor_options: 
  chunk_output_type: console
---

*Due Thursday, 04 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

40 points total.

Submit a single knitr file (named `homework7.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework7.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**21 points**

Use the following code to generate data for patients with repeated measures of A1C (a test for levels of blood glucose).

```{r}
genData <- function(n) {
    if(exists(".Random.seed", envir = .GlobalEnv)) {
        save.seed <- get(".Random.seed", envir= .GlobalEnv)
        on.exit(assign(".Random.seed", save.seed, envir = .GlobalEnv))
    } else {
        on.exit(rm(".Random.seed", envir = .GlobalEnv))
    }
    set.seed(n)
    subj <- ceiling(n / 10)
    id <- sample(subj, n, replace=TRUE)
    times <- as.integer(difftime(as.POSIXct("2005-01-01"),
                                 as.POSIXct("2000-01-01"), units='secs'))
    dt <- as.POSIXct(sample(times, n), origin='2000-01-01')
    mu <- runif(subj, 4, 10)
    a1c <- unsplit(mapply(rnorm, tabulate(id), mu, SIMPLIFY=FALSE), id)
    data.frame(id, dt, a1c)
}
x <- genData(500)
```

Perform the following manipulations: (3 points each)

1. Order the data set by `id` and `dt`.

```{r}
x1 = x[order(x$id,x$dt),]
```


2. For each `id`, determine if there is more than a one year gap in between observations.  Add a new row at the one year mark, with the `a1c` value set to missing.  A two year gap would require two new rows, and so forth.

```{r}
library(lubridate)

addmissing = function(y){
  delta = as.numeric(difftime(tail(y$dt, -1),
                   head(y$dt, -1),units = "days"))
  #print(delta)
  # compute the number of gaps for each time difference
  ngap = as.numeric(delta)%/%365
  # get the row index for the gaps
  rowidx_gap = seq_along(ngap)[ngap!=0]
  
  for (i in rowidx_gap){
    # get row i
    row_i = unlist(y[i,]) #here, the datetime will be coerced to UNIX form
    new_rows = data.frame(matrix(rep(row_i,ngap[i]),
                      byrow=T,nrow=ngap[i]))
    names(new_rows) = names(y)
    new_rows$dt = as.POSIXct(new_rows$dt, origin="1970-01-01")
    new_rows$id = as.integer(new_rows$id)
    for (j in 1:ngap[i]) new_rows$dt[j]=new_rows$dt[j]+years(j)
    new_rows$a1c = NA
    y = rbind(y,new_rows)
  }
  y[order(y$dt),]
}

datById = split(x1,x1$id) 
x2 = lapply(datById, addmissing)
#x2
```


3. Create a new column `visit`.  For each `id`, add the visit number.  This should be 1 to `n` where `n` is the number of observations for an individual.  This should include the observations created with missing a1c values.

```{r}
x3 = lapply(x2, function(y){
  y$visit = 1:nrow(y)
  y})
#x3
```


4. For each `id`, replace missing values with the mean `a1c` value for that individual.

```{r}
x4 = lapply(x3, function(y){
  m = mean(y$a1c,na.rm = T)
  y$a1c[is.na(y$a1c)] = m
  y
})
#x4
```


5. Print mean `a1c` for each `id`.

```{r}
x_final = do.call(rbind.data.frame, x4)
# mean `a1c` for each `id`
tapply(x_final$a1c,x_final$id,mean)
```

6. Print total number of visits for each `id`.

```{r}
tapply(x_final$visit,x_final$id,length)
```

7. Print the observations for `id = 15`.

```{r}
x4$`15`
```

### Question 2 ###

**16 points**

Install the `lexicon` package. Load the `sw_fry_1000` vector, which contains 1,000 common words.

```{r}
library(lexicon)
data('sw_fry_1000', package = 'lexicon')
#head(sw_fry_1000)
```

1. Remove all non-alphabetical characters and make all characters lowercase. Save the result as `a`.

```{r}
a = tolower(gsub('[^a-zA-Z]','',sw_fry_1000))
```

Use vector `a` for the following questions. (2 points each)

2. How many words contain the string "ar"?

```{r}
#grep('ar',sw_fry_1000,value = T)
length(grep('ar', a))
```

3. Find a six-letter word that starts with "l" and ends with "r".

```{r}
grep('^l[a-z]{4}r$', a, value = T)
```

4. Return all words that start with "col" or end with "eck".

```{r}
grep('^col|eck$', a, value = T)
```

5. Find the number of words that contain 4 or more adjacent consonants. Assume "y" is always a consonant.

```{r}
length(grep('[^aeiou]{4,}', a, value = T))
```

6. Return all words with a "q" that isn't followed by a "ui".

```{r}
# if q then not u => not qui
# if qu => then not i => not qui
# or ends with q or qu
grep('q[^u]|qu[^i]|qu$|q$', a, value = T)
```

7. Find all words that contain a "k" followed by another letter. Run the `table` command on the first character following the first "k" of each word.

```{r}
ks = grep('k[a-z]', a, value = T)
ks
table(substr(sub('[^k]*k',"",ks),1,1))
```

8. Remove all vowels. How many character strings are found exactly once?

```{r}
novowels = gsub('[aeiou]','',a)
#novowels
sum(table(novowels)==1)
```

### Question 3 ###

**3 points**

The first argument to most functions that fit linear models are formulas.  The following example defines the response variable `death` and allows the model to incorporate all other variables as terms. `.` is used to mean all columns not otherwise in the formula.

```{r}
url <- "https://raw.githubusercontent.com/couthcommander/Bios6301/main/datasets/haart.csv"
haart_df <- read.csv(url)[,c('death','weight','hemoglobin','cd4baseline')]
coef(summary(glm(death ~ ., data=haart_df, family=binomial(logit))))
```

Now imagine running the above several times, but with a different response and data set each time.  Here's a function:

```{r}
myfun <- function(dat, response) {
  form <- as.formula(response ~ .)
  coef(summary(glm(form, data=dat, family=binomial(logit))))
}
```

Unfortunately, it doesn't work. `tryCatch` is "catching" the error so that this file can be knit to PDF.

```{r}
tryCatch(myfun(haart_df, death), error = function(e) e)
```

What do you think is going on?  Consider using `debug` to trace the problem.

      In "as.formula(response ~ .)", the response is fixed as "response", 
      but what actually needed here is the name of our real response, i.e., death.


```{r,eval=F}
debugonce(myfun)
myfun(haart_df, death)
```

**5 bonus points**

Create a working function.

```{r}
myfun <- function(dat, response) { # response should be a character
  form <- as.formula(paste(response,'~ .'))
  coef(summary(glm(form, data=dat, family=binomial(logit))))
}
myfun(haart_df, "death")
```