---
title: 'Bios 6301: Assignment 8'
output: pdf_document
author: 'Lan Shi'
---

*Due Tuesday, 16 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

30 points total.

Submit a single knitr file (named `homework8.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework8.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

Install the `readxl` package and run the following

```{r}
library(readxl)
fn <- 'icd10.xlsx'
if(file.access(fn, mode = 4) == -1) {
    url <- "https://www.cdc.gov/nhsn/xls/icd10-pcs-pcm-nhsn-opc.xlsx"
    download.file(url, destfile = fn, mode = 'wb')
}
dat <- readxl::read_excel(fn, sheet = 2)
```

1. Show the class of `dat`. (1 point)

```{r}
class(dat)
```

2. Show the methods available for objects of the given class (if there are multiple classes, show methods for all classes). (3 points)

```{r}
methods(,class(dat)[1]) 
methods(,class(dat)[2]) 
methods(,class(dat)[3]) 
```

3. If you call `print(dat)`, what print method is being dispatched? (1 point)

```
print.tbl_df
```

4. Set the class of `dat` to be a data.frame. (1 point)

```{r}
class(dat) = 'data.frame'
```

5. If you call `print(dat)` again, what print method is being dispatched? (1 point)

```
print.data.frame
```

Define a new generic function `nUnique` with the code below.

```{r}
nUnique <- function(x) {
    UseMethod('nUnique')
}
```

6. Write a default method for `nUnique` to count the number of unique values in an element. (2 points)

```{r}
nUnique.default <- function(x) {
  length(unique(x))
}
```

7. Check your function (2 points)

```{r}
nUnique(letters) # should return 26
nUnique(sample(10, 100, replace = TRUE)) # should return 10 (probably)
```

8. Write a data.frame method for `nUnique` to operate on data.frame objects.
This version should return counts for each column in a data.frame. (2 points)

```{r}
nUnique.data.frame <- function(x) {
  sapply(x,\(y) length(unique(y)))
}
```

9. Check your function (2 points)

```{r}
nUnique(dat)
```

### Question 2 ###

**15 points**

Programming with classes.  The following function will generate random patient information.

```{r}
makePatient <- function() {
  vowel <- grep("[aeiou]", letters)
  cons <- grep("[^aeiou]", letters)
  name <- paste(sample(LETTERS[cons], 1), sample(letters[vowel], 1), sample(letters[cons], 1), sep='')
  gender <- factor(sample(0:1, 1), levels=0:1, labels=c('female','male'))
  dob <- as.Date(sample(7500, 1), origin="1970-01-01")
  n <- sample(6, 1)
  doa <- as.Date(sample(1500, n), origin="2010-01-01")
  pulse <- round(rnorm(n, 80, 10))
  temp <- round(rnorm(n, 98.4, 0.3), 2)
  fluid <- round(runif(n), 2)
  list(name, gender, dob, doa, pulse, temp, fluid)
}
```

1. Create an S3 class `medicalRecord` for objects that are a list with the named elements `name`, `gender`, `date_of_birth`, `date_of_admission`, `pulse`, `temperature`, `fluid_intake`. Note that an individual patient may have multiple measurements for some measurements.  Set the RNG seed to `8` and create a medical record by taking the output of `makePatient`.  Print the medical record, and print the class of the medical record. (5 points)

```{r}
set.seed(8)
mr = makePatient()
names(mr) = c('name', 'gender', 'date_of_birth', 'date_of_admission', 
              'pulse', 'temperature', 'fluid_intake')
class(mr) = 'medicalRecord'
print(mr)
print(class(mr))
```

2. Write a `medicalRecord` method for the generic function `mean`, which returns averages for pulse, temperature and fluids. Also write a `medicalRecord` method for `print`, which employs some nice formatting, perhaps arranging measurements by date, and `plot`, that generates a composite plot of measurements over time.  Call each function for the medical record created in part 1. (5 points)

```{r}
mean.medicalRecord <- function(mr){
    y = c(mean(mr$pulse),mean(mr$temperature),mean(mr$fluid_intake))
    names(y) = c('pulse', 'temperature', 'fluid_intake')
    y
}
mean(mr)

#arranging measurements by date, in a decreasing order
print.medicalRecord <- function(mr){
    y = do.call(cbind.data.frame, mr)
    y.arrange = y[order(y$date_of_admission,decreasing = T),]
    print(y.arrange)
}
print(mr)

# composite plot of measurements over time.
library(tidyr)
library(ggplot2)
plot.medicalRecord <- function(mr){
    y = do.call(cbind.data.frame, mr) %>% as_tibble %>%
      pivot_longer(5:7,names_to = 'type',values_to = "measurements")
    p = ggplot(y,aes(x=date_of_admission, 
                     y=measurements,color=type)) +
      geom_point(aes(shape=type)) + 
      geom_line() + theme_bw()  + 
      facet_grid(type ~ .,scales = "free")
    print(p)
}
plot(mr)
```

3. Create a further class for a cohort (group) of patients, and write methods for `mean` and `print` which, when applied to a cohort, apply mean or print to each patient contained in the cohort. Hint: think of this as a "container" for patients.  Reset the RNG seed to 8 and create a cohort of ten patients, then show the output for `mean` and `print`. (5 points)

```{r}
set.seed(8)
cohort = list()
name = NULL
for (i in 1:10){
  mr = makePatient()
  names(mr) = c('name', 'gender', 'date_of_birth', 
                'date_of_admission','pulse', 'temperature', 'fluid_intake')
  name = c(name,mr$name)
  cohort = c(cohort,list(mr))
}
names(cohort) = name
class(cohort) = 'cohort'
cohort[[5]]
mean.cohort = function(cohort){
  lapply(cohort,function(x){
    class(x) ='medicalRecord'
    mean(x)
  })
}
mean(cohort)

print.cohort <- function(cohort){
    invisible(lapply(cohort,function(x){
      class(x) ='medicalRecord'
      y = print(x)
    }))
}
print(cohort)
```
